---
title: "Making many heatgraphs"
author: "J.Norris NPS SCPN"
date: "November 2, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Set the path pathMain below 
### For detailed steps in Rstudio see table of contents.
### START CLEAN, remove all objects and detach packages ####

```{r startClean}

rm(list=ls())

## Detach non-base packages ####
detachNonBasePackages <- function() {
  ipack <- installed.packages()
  packListDf <-  as.data.frame(ipack[ ipack[,"Priority"] %in% c("base","recommended"), c("Package", "Priority")], stringsAsFactors = FALSE)
  baseOnly <- packListDf[which(packListDf$Priority == "base"),"Package"]
  baseOnlyFullName <- paste0("package:",packListDf[which(packListDf$Priority == "base"),"Package"])
  package.listAll <- search()[ifelse(unlist(gregexpr("package:",search()))==1,TRUE,FALSE)]
  package.listMost <- setdiff(package.listAll,baseOnlyFullName)
  if (length(package.listMost)>0)  for (package in package.listMost) detach(package, character.only=TRUE)
}
detachNonBasePackages()

```

### Load the libraries and set the paths and start and end years
```{r c1}
library(ggplot2)
library(gtable)
library(grid)
library(dplyr)


pathMain <- "Y:/Personal_Folders/Jodi/DataRequest_old/NazrugPresentation/ForGit/"
pathOutHeatGraphs <- paste0(pathMain,"outgraphs/")

load(paste0(pathMain,"ecoList2.Rda"))
load(paste0(pathMain,"uplandLU.Rda"))
load(paste0(pathMain,"ecositesNdviNoSnowScr.Rda"))
load(paste0(pathMain,"NdviDateJulOctMaxA.Rda"))
load(paste0(pathMain,"NdviDateMarJunMaxA.Rda"))
load(paste0(pathMain,"NdviDateAnnMaxA.Rda"))
load(paste0(pathMain,"NdviGpYearlyABCD.Rda"))
load(paste0(pathMain,"SnowEs2001_2014.Rda"))

yrStart <- 2001
yrEnd <- 2014
```
#
### BEGIN LOOP through ecosites for graphing ####
```{r ecoloop}
for (i in 1:nrow(ecoList2)) {
#i <- 3
ecoI <- ecoList2[i,"fireDraftEcoclassid"]
ecoIParkEco <- paste0(ecoList2[i,"Park"],ecoList2[i,"EcositeL"])
ecoLongNameR <- paste0(uplandLU[which(uplandLU$fireDraftEcoclassid == ecoI),"Park"],"  \n for the ", uplandLU[which(uplandLU$fireDraftEcoclassid == ecoI),"ecoclassname254"]," ecological site")    
ecoLongName <- paste0(ecoLongNameR[1])
# Get subsets and values based on this loop ####
myFilename <- paste0(pathOutHeatGraphs,ecoIParkEco,".jpg")
tempDf <- ecositesNdviNoSnowScr[which(ecositesNdviNoSnowScr$ecoclassid %in% ecoI),]
lowLimit0 <- quantile(tempDf$NdviMed, 0.03) 
lowLimit <- lowLimit0[[1]]
tempDf$NdviMod <- ifelse(tempDf$NdviMed < lowLimit,lowLimit,tempDf$NdviMed ) 
tempDfDateMonsPeak <- NdviDateJulOctMaxA[which(NdviDateJulOctMaxA$ecoclassid %in% ecoI),]
tempDfDateSprPeak <- NdviDateMarJunMaxA[which(NdviDateMarJunMaxA$ecoclassid %in% ecoI),]
tempDfDateAnnPeak <- NdviDateAnnMaxA[which(NdviDateAnnMaxA$ecoclassid %in% ecoI),]
tempDfAnnStats <- NdviGpYearlyABCD[which(NdviGpYearlyABCD$ecoclassid %in% ecoI),]
tempDFSnowDaily1C <- SnowEs2001_2014[which(SnowEs2001_2014$ecoclassid %in% ecoI & SnowEs2001_2014$SnowFlagSingCell == 1),]
tempDFSnowDaily2Pct <- SnowEs2001_2014[which(SnowEs2001_2014$ecoclassid %in% ecoI & SnowEs2001_2014$SnowPct > 0.02),]

avgAnnMin <- mean(tempDfAnnStats$NdviMinAnnP12) # what is the avg annual Ndvi minimum (screened to remove extremes)
avgAnnMax <- mean(tempDfAnnStats$NdviMaxAnn) # what is the avg annual Ndvi minimum (screened to remove extremes)
avgAnnAmp <- avgAnnMax- avgAnnMin # what is the average annual NDVI amplitude?
uLimit <- quantile(tempDf$NdviMed, 1)
yelBnd <- ((avgAnnMin + 0.1 * avgAnnAmp) - lowLimit) / (uLimit - lowLimit)  #color break for yellow
grnBnd <- ((avgAnnMin + 0.8 * avgAnnAmp) - lowLimit) / (uLimit - lowLimit) # color break for green
drkGrnBnd <- ((avgAnnMax + 0.2 * avgAnnAmp) - lowLimit) / (uLimit - lowLimit) # color break for dark green

# values for for color breaks
colBreaks <- c(0,yelBnd,grnBnd,drkGrnBnd,1)

## convert things below the lower limit to the limit itself, we'll mark the legend with 
## a less-than label
tempDf$NdviMod <- ifelse(tempDf$NdviMed < lowLimit,lowLimit,tempDf$NdviMed ) 

## MAKE HEATMAP with a legend ####
pMainLeg <-   ggplot(data = tempDf, aes(doyN, yearV)) + 
  geom_tile(aes(fill = NdviMod)) + 
  geom_point(data = tempDfDateMonsPeak, aes(x = doyN, y = yearV),pch = 2, size = 3, stroke = 1.5) + 
  geom_point(data = tempDfDateSprPeak, aes(x = doyN, y = yearV),pch = 6, size = 3, stroke = 1.5) +
  geom_point(data = tempDfDateAnnPeak, aes(x = doyN, y = yearV), pch = 1, size = 6, stroke = 1.5) + 
  scale_fill_gradientn(colours = c("purple",
                                  "yellow",
                                  "green",
                                  "dark green"), 
                       limits=c(lowLimit,uLimit),
                       values = colBreaks
                       ) + 
  scale_y_continuous("Year", limits = c(2000.5,2014.5), breaks = seq(yrStart,yrEnd), expand=c(0,0)) +
  scale_x_continuous("Month", 
                     limits = c(0,366), 
                     expand=c(0,0), 
                     breaks = c(1,32,60,91,121,152,182,213,244,274,305,335), 
                     labels = c("Jan-1","Feb-1","Mar-1","Apr-1","May-1","Jun-1","Jul-1","Aug-1","Sep-1","Oct-1","Nov-1","Dec-1")) +
  theme(legend.position = "left",
        legend.direction = "horizontal",
        legend.text = element_text(size = 0, color = "white"),
        axis.text.y = element_text( size=34),
       axis.text.x=element_text( size=34, angle = 90),
       axis.title.x = element_text(size=34),
       axis.line = element_line(colour="black", size=1),
       axis.title.y = element_blank(),
       axis.ticks.x = element_line(),
       axis.ticks.y =  element_blank(),
       axis.ticks.length = unit(0.2, "cm"),
        #                    panel.grid.major = element_blank(),
        #                    panel.grid.minor = element_blank(),
        panel.background = element_blank(),
       panel.border = element_rect(colour = "black", fill=NA, size=2)) +
  labs(fill="")


# ADD SNOW DATA PLOT#### 
pMainLeg <- pMainLeg + geom_tile(data=tempDFSnowDaily2Pct,aes(x=doyN,y=yearV),fill = "dark gray")
pMain <- pMainLeg + guides(fill = FALSE)  

## MAKE MARGINAL PLOT of annual stats####
pMargLeg <- ggplot(data = tempDfAnnStats, aes(x=yearV,y=NdviMaxJulOct, color = "Annual Stats")) +
  geom_line(color="blue",size=1.5) +  # monsoon line
  geom_point(data = tempDfAnnStats, aes(x=yearV,y=NdviMaxAnn,color="Annual peak"),size = 8, pch = 1, stroke = 1.5) + # annual line
  geom_point(aes(color = "Monsoon peak (JASO)"),fill = "blue", size = 4, pch = 24) + # monsoon symbol blue
  geom_line(data = tempDfAnnStats, aes(x=yearV,y=NdviMaxMarJun),color="red",size = 1.5) + # spring line
  geom_point(data = tempDfAnnStats, aes(x=yearV,y=NdviMaxMarJun,color="Spring Peak (MAMJ)"),fill = "red",size = 4, pch = 25) + # spring symbol
  geom_line(data = tempDfAnnStats, aes(x=yearV,y=NdviMinAnnP12),color="black",size = 1.5) + # ann min line
  geom_point(data = tempDfAnnStats, aes(x=yearV,y=NdviMinAnnP12,color="Annual base"),size = 4, pch = 19) + # ann min symbol                
  coord_flip() +
                    theme(axis.text.y = element_blank(),
                    axis.text.x=element_text(angle = 90,vjust = 0.5,size=34),
                    axis.title.x = element_text(size=34),
                    axis.line = element_line(colour="black", size=1),
                    axis.title.y = element_blank(),
                    axis.ticks.y =  element_blank(),
                    axis.ticks.length = unit(0.2, "cm"),
#                    panel.grid.major = element_blank(),
#                    panel.grid.minor = element_blank(),
                    panel.background = element_blank(),
                    legend.text=element_text(size=15),
                    legend.title=element_text(size=15)) +
                    scale_colour_manual(values = c("black","black","blue","red")) +
                    guides(color=guide_legend(override.aes=list(shape=c(19,1,24,25),fill = c("black","white","blue","red"), size = 5))) +
              scale_y_continuous(name ="Ndvi") + 
              labs(color = "Annual Statistics")

# same plot but without the legend so we can maniplate in the table
pMarg <- pMargLeg + guides(fill = FALSE, color = FALSE)  

## Combine the main heat plot and the marginal plot
gtMain <- ggplot_gtable(ggplot_build(pMain)) # this is for the main heatplot
gtMainLeg <- ggplot_gtable(ggplot_build(pMainLeg)) # this one is just to harvest the legend
gtMarg <- ggplot_gtable(ggplot_build(pMarg))
gtMargLeg <- ggplot_gtable(ggplot_build(pMargLeg)) # this one is just to harvest the legend
gtTitle <- textGrob(paste0("Smoothed NDVI values by day of year for 2001-2014 for ", ecoLongName),gp=gpar(fontsize=20,font=3))  
gtTitle2 <- textGrob("provisional data, values not finalized",gp=gpar(fontsize=18,font=3)) 
gtTitle3 <- textGrob(paste0("Less     More", "\ngreen   green"),gp=gpar(fontsize=20,font=3))  

#get those legends
leg1 <- gtMainLeg$grobs[[which(gtMainLeg$layout$name == "guide-box")]]
leg2 <- gtMargLeg$grobs[[which(gtMargLeg$layout$name == "guide-box")]]

# Get maximum widths and heights
maxWidth <- unit.pmax(gtMain$widths[2:3])
maxHeight <- unit.pmax(gtMain$heights[4:5], gtMarg$heights[4:5])


# Set the maximums in the gtables for gt1, gt2 and gt3 ####
gtMain$widths[2:3] <- as.list(maxWidth)

gtMain$heights[4:5] <- as.list(maxHeight)
gtMarg$heights[4:5] <- as.list(maxHeight)
# OPEN jpeg file ####
jpeg (myFilename,units = "mm",res=300,width = 430,height = 300)
gtFin <- gtable(widths = unit(c(7, 1.2,1.7), "null"), height = unit(c(0.75,0.30, 7), "null"))
#gtable  
# Instert gt1, gt2 and gt3 into the new gtable ####  position is row, column starting from the top left  
gtFin <- gtable_add_grob(gtFin, gtMain, 3, 1) # main heatmap
gtFin <- gtable_add_grob(gtFin, gtMarg, 3,2) # marginal plot
gtFin <- gtable_add_grob(gtFin, leg1, 2,2) # main legend
gtFin <- gtable_add_grob(gtFin, gtTitle, 1,1) # main title
gtFin <- gtable_add_grob(gtFin, gtTitle2, 2,1) # provisional data
gtFin <- gtable_add_grob(gtFin, leg2, 3,3) # marginal legend
gtFin <- gtable_add_grob(gtFin, gtTitle3, 1,2) # more/less green 

# And render the plot

grid.newpage()
grid.draw(gtFin)
dev.off()

}

# END of LOOP ####
```

### now let's just do the loope where i = 5 to show what's going on

##### setup data  for graphing ####
```{r JustLoop5}
i <- 5
ecoI <- ecoList2[i,"fireDraftEcoclassid"]
ecoIParkEco <- paste0(ecoList2[i,"Park"],ecoList2[i,"EcositeL"])
ecoLongNameR <- paste0(uplandLU[which(uplandLU$fireDraftEcoclassid == ecoI),"Park"],"  \n for the ", uplandLU[which(uplandLU$fireDraftEcoclassid == ecoI),"ecoclassname254"]," ecological site")    
ecoLongName <- paste0(ecoLongNameR[1])
# Get subsets and values based on this loop ####
myFilename <- paste0(pathOutHeatGraphs,ecoIParkEco,".jpg")
tempDf <- ecositesNdviNoSnowScr[which(ecositesNdviNoSnowScr$ecoclassid %in% ecoI),]
lowLimit0 <- quantile(tempDf$NdviMed, 0.03) 
lowLimit <- lowLimit0[[1]]
tempDf$NdviMod <- ifelse(tempDf$NdviMed < lowLimit,lowLimit,tempDf$NdviMed ) 
tempDfDateMonsPeak <- NdviDateJulOctMaxA[which(NdviDateJulOctMaxA$ecoclassid %in% ecoI),]
tempDfDateSprPeak <- NdviDateMarJunMaxA[which(NdviDateMarJunMaxA$ecoclassid %in% ecoI),]
tempDfDateAnnPeak <- NdviDateAnnMaxA[which(NdviDateAnnMaxA$ecoclassid %in% ecoI),]
tempDfAnnStats <- NdviGpYearlyABCD[which(NdviGpYearlyABCD$ecoclassid %in% ecoI),]
tempDFSnowDaily1C <- SnowEs2001_2014[which(SnowEs2001_2014$ecoclassid %in% ecoI & SnowEs2001_2014$SnowFlagSingCell == 1),]
tempDFSnowDaily2Pct <- SnowEs2001_2014[which(SnowEs2001_2014$ecoclassid %in% ecoI & SnowEs2001_2014$SnowPct > 0.02),]

avgAnnMin <- mean(tempDfAnnStats$NdviMinAnnP12) # what is the avg annual Ndvi minimum (screened to remove extremes)
avgAnnMax <- mean(tempDfAnnStats$NdviMaxAnn) # what is the avg annual Ndvi minimum (screened to remove extremes)
avgAnnAmp <- avgAnnMax- avgAnnMin # what is the average annual NDVI amplitude?
uLimit <- quantile(tempDf$NdviMed, 1)
yelBnd <- ((avgAnnMin + 0.1 * avgAnnAmp) - lowLimit) / (uLimit - lowLimit)  #color break for yellow
grnBnd <- ((avgAnnMin + 0.8 * avgAnnAmp) - lowLimit) / (uLimit - lowLimit) # color break for green
drkGrnBnd <- ((avgAnnMax + 0.2 * avgAnnAmp) - lowLimit) / (uLimit - lowLimit) # color break for dark green

# values for for color breaks
colBreaks <- c(0,yelBnd,grnBnd,drkGrnBnd,1)

## convert things below the lower limit to the limit itself, we'll mark the legend with 
## a less-than label
tempDf$NdviMod <- ifelse(tempDf$NdviMed < lowLimit,lowLimit,tempDf$NdviMed ) 
```
#### MAKE HEATMAP with a legend ####

```{r EC1}
pMainLeg <-   ggplot(data = tempDf, aes(doyN, yearV)) + 
  geom_tile(aes(fill = NdviMod)) + 
  geom_point(data = tempDfDateMonsPeak, aes(x = doyN, y = yearV),pch = 2, size = 3, stroke = 1.5) + 
  geom_point(data = tempDfDateSprPeak, aes(x = doyN, y = yearV),pch = 6, size = 3, stroke = 1.5) +
  geom_point(data = tempDfDateAnnPeak, aes(x = doyN, y = yearV), pch = 1, size = 6, stroke = 1.5) + 
  scale_fill_gradientn(colours = c("purple",
                                  "yellow",
                                  "green",
                                  "dark green"), 
                       limits=c(lowLimit,uLimit),
                       values = colBreaks
                       ) + 
  scale_y_continuous("Year", limits = c(2000.5,2014.5), breaks = seq(yrStart,yrEnd), expand=c(0,0)) +
  scale_x_continuous("Month", 
                     limits = c(0,366), 
                     expand=c(0,0), 
                     breaks = c(1,32,60,91,121,152,182,213,244,274,305,335), 
                     labels = c("Jan-1","Feb-1","Mar-1","Apr-1","May-1","Jun-1","Jul-1","Aug-1","Sep-1","Oct-1","Nov-1","Dec-1")) +
  theme(legend.position = "left",
        legend.direction = "horizontal",
        legend.text = element_text(size = 0, color = "white"),
        axis.text.y = element_text( size=34),
       axis.text.x=element_text( size=34, angle = 90),
       axis.title.x = element_text(size=34),
       axis.line = element_line(colour="black", size=1),
       axis.title.y = element_blank(),
       axis.ticks.x = element_line(),
       axis.ticks.y =  element_blank(),
       axis.ticks.length = unit(0.2, "cm"),
        #                    panel.grid.major = element_blank(),
        #                    panel.grid.minor = element_blank(),
        panel.background = element_blank(),
       panel.border = element_rect(colour = "black", fill=NA, size=2)) +
  labs(fill="")

# ADD SNOW DATA PLOT#### 
pMainLeg <- pMainLeg + geom_tile(data=tempDFSnowDaily2Pct,aes(x=doyN,y=yearV),fill = "dark gray")
```

#### display the plot for showNtell
```{r EC2}
pMainLeg
```

#### same plot but without the legend ####
#### so we can combine it with the marginal plot

```{r EC3}
pMain <- pMainLeg + guides(fill = FALSE)  
# display the plot for showNtell
pMain
```

#### MAKE MARGINAL PLOT of annual stats####
```{r EC4}
pMargLeg <- ggplot(data = tempDfAnnStats, aes(x=yearV,y=NdviMaxJulOct, color = "Annual Stats")) +
  geom_line(color="blue",size=1.5) +  # monsoon line
  geom_point(data = tempDfAnnStats, aes(x=yearV,y=NdviMaxAnn,color="Annual peak"),size = 8, pch = 1, stroke = 1.5) + # annual line
  geom_point(aes(color = "Monsoon peak (JASO)"),fill = "blue", size = 4, pch = 24) + # monsoon symbol blue
  geom_line(data = tempDfAnnStats, aes(x=yearV,y=NdviMaxMarJun),color="red",size = 1.5) + # spring line
  geom_point(data = tempDfAnnStats, aes(x=yearV,y=NdviMaxMarJun,color="Spring Peak (MAMJ)"),fill = "red",size = 4, pch = 25) + # spring symbol
  geom_line(data = tempDfAnnStats, aes(x=yearV,y=NdviMinAnnP12),color="black",size = 1.5) + # ann min line
  geom_point(data = tempDfAnnStats, aes(x=yearV,y=NdviMinAnnP12,color="Annual base"),size = 4, pch = 19) + # ann min symbol                
  coord_flip() +
                    theme(axis.text.y = element_blank(),
                    axis.text.x=element_text(angle = 90,vjust = 0.5,size=34),
                    axis.title.x = element_text(size=34),
                    axis.line = element_line(colour="black", size=1),
                    axis.title.y = element_blank(),
                    axis.ticks.y =  element_blank(),
                    axis.ticks.length = unit(0.2, "cm"),
#                    panel.grid.major = element_blank(),
#                    panel.grid.minor = element_blank(),
                    panel.background = element_blank(),
                    legend.text=element_text(size=15),
                    legend.title=element_text(size=15)) +
                    scale_colour_manual(values = c("black","black","blue","red")) +
                    guides(color=guide_legend(override.aes=list(shape=c(19,1,24,25),fill = c("black","white","blue","red"), size = 5))) +
              scale_y_continuous(name ="Ndvi") + 
              labs(color = "Annual Statistics")
```

#### display the plot for showNtell
```{r EC5}
pMargLeg
```

#### same plot but without the legend so we can maniplate in the table
```{r EC6}
pMarg <- pMargLeg + guides(fill = FALSE, color = FALSE)  
# display the plot for showNtell
pMarg
```

#### Start building and creating text objects
```{r EC7}
gtMain <- ggplot_gtable(ggplot_build(pMain)) # this is for the main heatplot
gtMainLeg <- ggplot_gtable(ggplot_build(pMainLeg)) # this one is just to harvest the legend
```

#### now lets do it using the alternate ggplotGrob function
  ##### I don't know what the difference is, but there may be different settings available
```{r EC8}
gtMarg <- ggplotGrob(pMarg)
gtMargLeg <- ggplotGrob(pMargLeg) # this one is just to harvest the legend
```
#### Lets make some title text objects 
```{r EC9}
gtTitle <- textGrob(paste0("Smoothed NDVI values by day of year for 2001-2014 for ", ecoLongName),gp=gpar(fontsize=20,font=3))  
gtTitle2 <- textGrob("provisional data, values not finalized",gp=gpar(fontsize=18,font=3)) 
gtTitle3 <- textGrob(paste0("Less     More", "\ngreen   green"),gp=gpar(fontsize=20,font=3))  
```

####get those legends
```{r EC10}
leg1 <- gtMainLeg$grobs[[which(gtMainLeg$layout$name == "guide-box")]]
leg2 <- gtMargLeg$grobs[[which(gtMargLeg$layout$name == "guide-box")]]
```

#### see what the oject looks like ####
```{r EC11}
gtMargLeg
```

#### OPEN jpeg file as display ####

```{r EC12}
jpeg (myFilename,units = "mm",res=300,width = 430,height = 300)
```

#### create our final gtable object, setting the relative row and column dimensions
```{r EC13}
gtFin <- gtable(widths = unit(c(7, 1.2,1.7), "null"), height = unit(c(0.75,0.30, 7), "null"))
```
#### Instert gt1, gt2 and gt3 into the new gtable ####
####  position is row, column starting from the top left  
```{r EC14}
gtFin <- gtable_add_grob(gtFin, gtMain, 3, 1) # main heatmap
gtFin <- gtable_add_grob(gtFin, gtMarg, 3,2) # marginal plot
gtFin <- gtable_add_grob(gtFin, leg1, 2,2) # main legend
gtFin <- gtable_add_grob(gtFin, gtTitle, 1,1) # main title
gtFin <- gtable_add_grob(gtFin, gtTitle2, 2,1) # provisional data
gtFin <- gtable_add_grob(gtFin, leg2, 3,3) # marginal legend
gtFin <- gtable_add_grob(gtFin, gtTitle3, 1,2) # more/less green 
```

#### And render the plot, it looks better in the saved files so check the outgraphs directory
```{r EC15}
grid.newpage()
grid.draw(gtFin)
dev.off()
grid.draw(gtFin)
```

